Squint: Lossy Hierarchical Compression on Symbol Streams

### 2012-12-11 tex/paper.tex
Added in prelim background section

In this paper we develop 2 algorithms and implement a system for lossily
compressing text while preserving its hierarchical structures. The goal
was to preserve the feel of a text at both micro and macro levels. 
Since grammar-based compression schemes were designed to capture
these structures, we built
Music Final Composition: Dec 10
our system on top of the existing \Sequitur grammar inference algorithm. 
Starting with the CFG generated by a lossless grammar
inference algorithm, we repeatedly simplify it, throwing out details
along the way to emphasize repetitions. Our main contribution lies in the
development of 2 algorithms for ``simplifying'' CFG grammars while 
preserving their structure.

@@

TODO

\subsection{Grammar Compression}

TODO

@@

TODO

Jpeg, Mpeg: using parts of previous frame to encode next frame, 
Fractal: Using self-similarity to encode

@@

The scope of what we shall call a \emph{lossifier} algorithm is to take
an admissible CFG and simplify it. The output CFG should be smaller
than the input CFG but generate a string which preserves the
structure in the original string as much as possible.

Building off the ideas in jpeg compression discussed
earlier, one idea for simplifying data is to look at a part
of the data and try to match it as closely as possible with an
existing set of primitives. In particular, just as with Mpeg and
fractal compression, to approximate one part of the grammar
we can use \emph{other parts of the grammar}. This is the seed for
both of our algorithms. Given a rule $S \rightarrow rhs$, we will
try to replace occurences of $S$ with other similar variables.

>>>>

In this paper we develop 2 algorithms and implement a system for lossily
compressing text while preserving its hierarchical structures. The goal
was to preserve the feel of a text at both micro and macro levels. 
Since grammar-based compression schemes were designed to capture
these structures, we built
our system on top of the existing \Sequitur grammar inference algorithm. 
Starting with the CFG generated by a lossless grammar
inference algorithm, we repeatedly simplify it, throwing out details
along the way to emphasize repetitions. Our main contribution lies in the
development, implementation, and evaluation of 2 algorithms for 
``simplifying'' CFG grammars while preserving their structure.

@@

The work in this paper builds upon existing research in lossless
grammar based compression, and also draws inspiration from many
techniques used in lossy image and video compression.

\subsection{Grammar Compression}
As described in the introduction, the proposed research shares many of its
goals with the \Sequitur algorithm described in \cite{sequitur}.  The authors
of \cite{sequitur} develop an effective algorithm based on iteratively
rewriting grammars to keep them small and efficient, and produce interesting
grammar-based analyses of texts and musical scores.

The same author in \cite{nevillphd}
explores a variety of ad-hoc extensions to the \Sequitur algorithm which
improve its compression performance on structured data.  These ranged from
introducing domain-specific constraints to their grammar, adding a few steps of
backtracking to their normally greedy grammar formation, to guessing
unifications in attempting to infer recursive grammar rules.  Though these may
invalidate some of their theoretical results on the asymptotic performance of
\Sequitur, in practice they seem to have worked and add nicely to the grammar
inference framework.

The intution that grammar inference can support many detailed policies is made
more formal in \cite{grammarcodes}.  The authors classify the properties a
grammar needs to function as a good compressor, and moreover give a set of
reduction rules for putting a grammar into an appropriate
``irreducible'' form.  Within this
context, grammar inference algorithms similar to \Sequitur can be described as
simple applications of their reduction rules to different ways of generating
base grammars.  Even more examples of the variety of grammar inference schemes
possible under this general framework of reducing grammars can be found in
\cite{efficientgreedy}

There has also been very interesting work analyzing how grammar based
algorithms compare with the theoretical best lossless grammar for a string, in
the line of \cite{approximation}.  However, these analytic bounds for lossless
grammars are outside the scope of our proposed research.

In summary, established work on grammar inference algorithms provides a solid
set of tools for experimenting with the kinds of local modifications we propose
to make to improve grammar structure by introducing lossiness.

@@

While we were unable to find work in the area of lossy grammar compression, a
number of existing techniques for lossy hierarchical compression more broadly
may be applicable.

Kieffer briefly identifies two primary methods for lossy hierarchical
compression: wavelet-based and fractal-based schemes \cite{tutorial}.  In
wavelet-based schemas, a signal is compressed by recording the largest
coefficients of its wavelet decomposition, in effect
trying to represent a signal as closely as possible using wavelet
building blocks. The JPEG image format uses a similar technique in
order to compress images lossily \cite{jpeg}. Fractal based compression
takes this a step further and represents parts of a signal using self-similarity,
encoding an image for instance as a fixed point of contraction maps. This is
similar in spirit to what is done in MPEG-1, where parts of one frame can be
transformed to encode parts of the next frame.

Work has also been done regarding the lossy compression of particular input
formats.  For text, methods such as replacing words with synonyms
\cite{semanticlossy} or reordering the middle letters of a word (i.e., all
letters save for the first and last) \cite{semilossless} have been explored.
However, these are very specific tricks to optimize text, and not more broadly
applicable to compressing hierarchical streams.

@@

In order to make our algorithm for lossily compressing streams
as general as possible, we decided
to approach the problem of lossy compression by starting with a lossless
grammar inference algorithm. 
By building off of a grammar inference algorithm, we
hope to take advantage of some of the partial structures
that are visible in a CFG representation of a stream. 
In this project, we chose \Sequitur, but
any way of obtaining a grammar could be used. Then, we would transform
the grammar, introducing loss and perform compressions by simplifying
the grammar. Thus our compression system is a multi-step process:

\begin{itemize}
\item Infer a CFG representation of a stream (possibly using Sequitur)
\item Simplify the CFG, possibly introducing loss
\item Output the CFG
\end{itemize}

Our core new contribution comes in step two, in what we shall call a \emph{lossifier} 
algorithm, i.e. an algorithm that takes
an admissible CFG and simplifies it. The output CFG should be smaller
than the input CFG but generate a string which preserves the
structure in the original string as much as possible.

Building off the ideas in lossy compression discussed
earlier, one idea for simplifying data is to look at a part
of the data and try to match it as closely as possible with an
existing set of primitives. In particular, just as with Mpeg and
fractal compression, to approximate one part of the grammar
we can use \emph{other parts of the grammar}. This is the seed for
both of our algorithms. Given a rule $S \rightarrow rhs$, we will
try to replace occurences of $S$ with other similar variables.
